---
title: "National phenology network"
format: gfm
---

This data set contains data from the [USA National Phenology Network](https://www.usanpn.org/about).

In R, we need 2 packages to download and visualize the data. First, check if the packages are already installed. Install them if they are not:

```{r}
packages <- c("tidyverse", "devtools") 
new.packages <- packages[!(packages %in% installed.packages()[,"Package"])] 
if(length(new.packages)>0) install.packages(new.packages)
```

Then, load them and install '[rnpn](https://rdrr.io/cran/rnpn/f/README.md)', an R client for interacting with the USA National Phenology Network data web services.

```{r}
lapply(packages, library, character.only = TRUE)
install_github("usa-npn/rnpn")
library(rnpn)
```

Download data about Woodhouse's toads:

```{r}
toads <- npn_download_status_data(request_source='Your Name',years = c(2000:2020), species_id=c(237))
```

Plot phenophases as a function of DOY per state:

```{r}
ggplot(toads) +
  geom_point(aes(day_of_year, phenophase_description)) +
  facet_wrap(.~state) +
  theme_bw() +
  xlab('Day of year') +
  ylab('Phenophase')
```
