---
title: "National phenology network"
author: "Ty Tuff, ESIIL Data Scientist"
date: "2023-05-21"
format: gfm
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}  
library(knitr)  
library(reticulate)  
knitr::knit_engines$set(python = reticulate::eng_python)  
knitr::opts_chunk$set(
  cache=TRUE, message=FALSE, warning=FALSE, results='hide')  
```

This data set contains data from the [USA National Phenology Network](https://www.usanpn.org/about).

The USA National Phenology Network (USA-NPN) is an organization that monitors the influence of climate on the phenology of plants, animals, and landscapes. Phenology is the study of cyclic and seasonal natural phenomena, especially in relation to climate and plant and animal life.

The USA-NPN brings together citizen scientists, government agencies, non-profit groups, educators, and students of all ages to monitor the impacts of climate change on plants and animals in the United States.

The data set collected by the USA-NPN, which you mentioned, contains a wide array of information about phenological trends and patterns across the country. This data can be critical in understanding the impact of climate change on various species and ecosystems. It provides invaluable information that helps scientists predict trends and potentially develop strategies for mitigating the effects of climate change on our environment.

The data set can include data on the timing of leaf-out, flowering, and fruiting of plants; the emergence of insects and the migration of birds; the peak of allergy season; the timing of agricultural crop stages; and many other biological events.


In R, we need 2 packages to download and visualize the data. First, check if the packages are already installed. Install them if they are not:

```{r, results='hide',cache=TRUE, warning=FALSE, messages=FALSE}
packages <- c("tidyverse", "devtools") 
new.packages <- packages[!(packages %in% installed.packages()[,"Package"])] 
if(length(new.packages)>0) install.packages(new.packages)
```

Then, load them and install '[rnpn](https://rdrr.io/cran/rnpn/f/README.md)', an R client for interacting with the USA National Phenology Network data web services.

```{r, results='hide',cache=TRUE, warning=FALSE, messages=FALSE}
lapply(packages, library, character.only = TRUE)
install_github("usa-npn/rnpn")
library(rnpn)
```

Download data about Woodhouse's toads:

```{r, cache=TRUE, warning=FALSE, messages=FALSE}
toads <- npn_download_status_data(request_source='Your Name',years = c(2000:2020), species_id=c(237))
```

Plot phenophases as a function of DOY per state:

```{r, cache=TRUE, warning=FALSE, messages=FALSE}
ggplot(toads) +
  geom_point(aes(day_of_year, phenophase_description)) +
  facet_wrap(.~state) +
  theme_bw() +
  xlab('Day of year') +
  ylab('Phenophase')
```
