---
title: "dryad"
author: "Ty Tuff, ESIIL Data Scientist"
date: "2023-05-21"
format: gfm
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}  
library(knitr)  
library(reticulate)  
knitr::knit_engines$set(python = reticulate::eng_python)  
knitr::opts_chunk$set(
  cache=TRUE, message=FALSE, warning=FALSE, results='hide')  
```

Dryad is a general-purpose open-access repository that makes the data underlying scientific publications discoverable, freely reusable, and citable. Hosted by the California Digital Library (CDL), Dryad has been a pioneer in the movement to make scientific research data openly available for researchers, students, and the public since its inception.

In the spirit of promoting open science, Dryad aims to preserve and provide access to data that underpins findings in peer-reviewed publications in the natural sciences, as well as certain areas of the social sciences and clinical studies. Its primary goal is to facilitate data reuse and new discoveries while giving proper credit to the data authors and ensuring data preservation.

The repository accepts data submissions associated with any peer-reviewed manuscript. It supports a wide variety of data types and formats, making it an extremely versatile resource for researchers from various fields. Dryad uniquely assigns Digital Object Identifiers (DOIs) to datasets to ensure that data can be correctly cited.

In addition to providing access to data, Dryad also provides a platform for researchers to comply with funder and journal data archiving policies. By promoting data archiving best practices, Dryad contributes to the broader movement of responsible data management in the scientific community.

## The phantom chorus
To access a dataset, you'll first need to choose a dataset and get its DOI (Digital Object Identifier) or a direct link to download the dataset. In this example, we'll use a dataset called "The phantom chorus: birdsong boosts human well-being in protected areas" by Dr. Clinton Francis with DOI: 10.5061/dryad.vmcvdncs3.

https://datadryad.org/stash/dataset/doi:10.5061/dryad.wwpzgmsgx

Citation

Francis, Clinton (2020), The phantom chorus: birdsong boosts human well-being in protected areas, Dryad, Dataset, https://doi.org/10.5061/dryad.wwpzgmsgx

Abstract

Spending time in nature is known to benefit human health and well-being, but evidence is mixed as to whether biodiversity or perceptions of biodiversity contribute to these benefits. Perhaps more importantly, little is known about the sensory modalities by which humans perceive biodiversity and obtain benefits from their interactions with nature. Here, we used a “phantom bird song chorus” consisting of hidden speakers to experimentally increase audible birdsong biodiversity during “on” and “off” (i.e., ambient conditions) blocks on two trails to study the role of audition in biodiversity perception and self-reported well-being among hikers. Hikers exposed to the phantom chorus reported higher levels of restorative effects compared to those that experienced ambient conditions on both trails; however, increased restorative effects were directly linked to the phantom chorus on one trail and indirectly linked to the phantom chorus on the other trail through perceptions of avian biodiversity. Our findings add to a growing body of evidence linking mental health to nature experiences and suggest that audition is an important modality by which natural environments confer restorative effects. Finally, our results suggest that maintaining or improving natural soundscapes within protected areas may be an important component to maximizing human experiences.

<iframe width="100%" 
height="700" 
src="https://doi.org/10.5061/dryad.wwpzgmsgx" 
style="border:none" 
title="Contributed by NEON">
</iframe>

R code:

```{r, eval=FALSE}
install.packages("rdryad")
install.packages("ggplot2")
```




```{r, cache=TRUE}
library(rdryad)
library(ggplot2)


search_results <- dryad_datasets()
print(search_results$data)

# Get the first DOI from the search results
doi <-  "https://doi.org/10.5061/dryad.wwpzgmsgx"
clean_doi <- gsub("https://doi.org/", "", doi)

files <- dryad_download(clean_doi)
print(files)
a <- unlist(files)
short_files <- as.data.frame(a)[,1]

# Identify the CSV file from the extracted files
csv_file <- list.files("/Users/ty/Library/Caches/R/rdryad/10_5061_dryad_wwpzgmsgx", pattern = ".csv", full.names = TRUE)[2]

# Read the CSV data
data <- read.csv(csv_file)

ggplot(data=data, aes(y=Perceivedrestoration, Sound_Comp)) +
  geom_point()
  
```



```{python, cache=TRUE, eval=FALSE}
import requests
import pandas as pd
import matplotlib.pyplot as plt
import io

# Search for datasets related to the term "habitat"
base_url = "https://datadryad.org/api/v2/datasets?"
params = {"query": "habitat"}

response = requests.get(base_url, params=params)
search_results = response.json()
print(search_results)

# Get the first DOI from the search results
doi = "https://doi.org/10.5061/dryad.wwpzgmsgx"
clean_doi = doi.replace("https://doi.org/", "")

# Download the dataset
files_url = f"https://datadryad.org/api/v2/datasets/{clean_doi}/download"
response = requests.get(files_url)
files = response.json()
print(files)

# Identify the CSV file from the extracted files
csv_file_url = None
for file in files['data']:
    if '.csv' in file['attributes']['name']:
        csv_file_url = file['links']['download']
        break

# Read the CSV data
response = requests.get(csv_file_url)
data = pd.read_csv(io.StringIO(response.text))

# Plot the data using pandas and matplotlib
# Replace 'Perceivedrestoration' and 'Sound_Comp' with the actual column names in your dataset
ax = data.plot(x='Perceivedrestoration', y='Sound_Comp', kind='scatter')
ax.set_xlabel('Perceivedrestoration')
ax.set_ylabel('Sound_Comp')
plt.show()
```




