---
title: "dryad"
format: gfm
---
Dryad is a repository of curated, open access, data files associated with published articles in the sciences and social sciences. To access a dataset, you'll first need to choose a dataset and get its DOI (Digital Object Identifier) or a direct link to download the dataset. In this example, we'll use a dataset called "The phantom chorus: birdsong boosts human well-being in protected areas" by Dr. Clinton Francis with DOI: 10.5061/dryad.vmcvdncs3.

https://datadryad.org/stash/dataset/doi:10.5061/dryad.wwpzgmsgx

R code:

```{r, eval=FALSE}
install.packages("rdryad")
install.packages("ggplot2")
```




```{r, cache=TRUE}
library(rdryad)
library(ggplot2)


search_results <- dryad_datasets()
print(search_results$data)

# Get the first DOI from the search results
doi <-  "https://doi.org/10.5061/dryad.wwpzgmsgx"
clean_doi <- gsub("https://doi.org/", "", doi)

files <- dryad_download(clean_doi)
print(files)
a <- unlist(files)
short_files <- as.data.frame(a)[,1]

# Identify the CSV file from the extracted files
csv_file <- list.files("/Users/ty/Library/Caches/R/rdryad/10_5061_dryad_wwpzgmsgx", pattern = ".csv", full.names = TRUE)[2]

# Read the CSV data
data <- read.csv(csv_file)

ggplot(data=data, aes(y=Perceivedrestoration, Sound_Comp)) +
  geom_point()
  
```



```{python, cache=TRUE, eval=FALSE}
import requests
import pandas as pd
import matplotlib.pyplot as plt
import io

# Search for datasets related to the term "habitat"
base_url = "https://datadryad.org/api/v2/datasets?"
params = {"query": "habitat"}

response = requests.get(base_url, params=params)
search_results = response.json()
print(search_results)

# Get the first DOI from the search results
doi = "https://doi.org/10.5061/dryad.wwpzgmsgx"
clean_doi = doi.replace("https://doi.org/", "")

# Download the dataset
files_url = f"https://datadryad.org/api/v2/datasets/{clean_doi}/download"
response = requests.get(files_url)
files = response.json()
print(files)

# Identify the CSV file from the extracted files
csv_file_url = None
for file in files['data']:
    if '.csv' in file['attributes']['name']:
        csv_file_url = file['links']['download']
        break

# Read the CSV data
response = requests.get(csv_file_url)
data = pd.read_csv(io.StringIO(response.text))

# Plot the data using pandas and matplotlib
# Replace 'Perceivedrestoration' and 'Sound_Comp' with the actual column names in your dataset
ax = data.plot(x='Perceivedrestoration', y='Sound_Comp', kind='scatter')
ax.set_xlabel('Perceivedrestoration')
ax.set_ylabel('Sound_Comp')
plt.show()
```




