---
title: "Institute of Museum and Library Services (IMLS) Data Catalog"
author: "Ty Tuff, ESIIL Data Scientist"
date: "2023-05-21"
format: gfm
editor_options: 
  markdown: 
    wrap: 72
---
```{r setup, include=FALSE}  
library(knitr)  
library(reticulate)  
knitr::knit_engines$set(python = reticulate::eng_python)  
```


The Institute of Museum and Library Services (IMLS) is a key source of data and research pertaining to libraries and museums in the United States. The IMLS Data Catalog, an invaluable tool for professionals, researchers, and policymakers, provides comprehensive datasets that offer deep insights into these cultural institutions and their impact on their communities.

<https://www.imls.gov/research-tools/data-collection>

A significant contribution within this catalog is the Public Libraries Survey (PLS) dataset. The PLS is a yearly national census of public libraries in the U.S., aiming to provide reliable and detailed information concerning the state of public libraries nationwide. This dataset collects and compiles data about crucial aspects of library operations, including library usage, staffing levels, operating finances, types of services offered, and collection sizes.

The PLS dataset serves as a vital tool for studying the librarian profession and understanding library practices across the country. For instance, it can help identify trends in library service delivery, evaluate the efficiency of different libraries, and assess the impact of budgetary changes on library operations and services.

Moreover, it enables comparison of libraries of different sizes, types, and geographical locations, assisting in benchmarking and best practice identification. It also provides researchers and policymakers with the data needed to make informed decisions about library funding, strategic planning, and policy development.

In summary, the IMLS Data Catalog, and particularly the PLS dataset, is a rich resource for anyone interested in understanding the current state of libraries in the United States, their operations, their challenges, and their contributions to their communities.

Here's how to download the Public Libraries Survey dataset and create a plot using R and Python:

R code:

```{r, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Load required libraries
library(tidyverse)
library(readr)
library(curl)

# Download the dataset
url <- "https://www.imls.gov/sites/default/files/2022-07/pls_fy2020_csv.zip"
temp_file <- tempfile(fileext = ".zip")
curl_download(url, temp_file)

# Unzip the dataset
temp_dir <- tempdir()
unzip(temp_file, exdir = temp_dir)

# Read the CSV file
csv_file <- list.files(temp_dir, pattern = "*.csv", full.names = TRUE)
data <- read_csv(csv_file[1]) # Corrected this line

# Create a plot (example: number of libraries by state)
state_counts <- data %>% count(STABR)
ggplot(state_counts, aes(x = reorder(STABR, -n), y = n)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Number of Libraries by State",
       x = "State",
       y = "Number of Libraries") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


Python code:

```{python, cache=TRUE, message=FALSE, warning=FALSE, results='hide', eval=FALSE}

import requests
import pandas as pd
import matplotlib.pyplot as plt
import zipfile
import os
from io import BytesIO

# Download the dataset
url = "https://www.imls.gov/sites/default/files/2022-07/pls_fy2020_csv.zip"
response = requests.get(url)

# Unzip the dataset
with zipfile.ZipFile(BytesIO(response.content)) as zip_ref:
    zip_ref.extractall()

# Find the CSV file
csv_file = next(file for file in os.listdir() if file.endswith(".csv"))

# Read the CSV file
data = pd.read_csv(csv_file)

# Count the number of libraries by state
state_counts = data['STABR'].value_counts()

# Create a plot
plt.figure(figsize=(10,8))
state_counts.sort_values().plot(kind = 'barh')
plt.title('Number of Libraries by State')
plt.xlabel('Number of Libraries')
plt.ylabel('State')
plt.grid(axis='x')
plt.show()
```

Both the R and Python code above download the Public Libraries Survey dataset, convert it to a data frame, and create a bar plot of the total library staff by state. Note that the dataset provided by the IMLS Data Catalog API might have a limit on the number of records returned, so you may need to adjust the API request to retrieve a more extensive dataset.
