---
title: "FIRED"
author: "Virginia Iglesias, ESIIL Data Scientist"
date: "2023-05-21"
format: gfm
editor_options: 
  markdown: 
    wrap: 72
---

The Fire Event Delineation (FIRED) data set is a significant contribution to the field of environmental science, specifically in the study and monitoring of wildfires. Covering an extensive period from November 2001 to January 2022, this data set provides comprehensive spatial information about wildfire events in the conterminous United States, making it an invaluable resource for researchers, policymakers, and environmental practitioners alike.

The FIRED product is unique as it offers both event-level and daily-level polygons of wildfire incidents. This dual-level detail allows for precise spatial and temporal mapping and tracking of wildfire occurrences. Event-level polygons encapsulate the entire spatial extent of individual fire events, whereas daily-level polygons provide day-to-day progression information for each fire, effectively illustrating the development and spread of fires over time.

This data set contains event- and daily-level polygons for the [Fire Event Delineation](https://scholar.colorado.edu/concern/datasets/d504rm74m) (FIRED) product for the conterminous US from November 2001 to January 2022. You can find more information [here](https://www.mdpi.com/2072-4292/12/21/3498).

In R, we need 3 packages to download and visualize the data. First, check if the packages are already installed. Install them if they are not:

```{r, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
packages <- c("tidyverse", "httr", "sf") 
new.packages <- packages[!(packages %in% installed.packages()[,"Package"])] 
if(length(new.packages)>0) install.packages(new.packages) 
```

Then, load them:

```{r, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
lapply(packages, library, character.only = TRUE)
```

Download the data set:

```{r, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
url <- "https://scholar.colorado.edu/downloads/zw12z650d" 
fired <- GET(url) 
data_file <-"fired.zip" 
writeBin(content(fired, "raw"), data_file)

# Unzip the file
unzip(data_file)
```

Read the data set:

```{r, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
fired <- st_read("fired_conus_ak_to_January_2022_gpkg_shp/conus_ak_to2022001_events.shp") 
```

Plot fire duration as a function of ignition day:

```{r, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
ggplot(fired) +
  geom_point(aes(ig_day, event_dur)) +
  theme_bw() +
  xlab('Day') +
  ylab('Event duration (days)')
```

In Python, we need 5 libraries to download and visualize the data.

```{python, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
import requests 
import zipfile 
import geopandas as gpd 
import matplotlib.pyplot as plt
import seaborn as sns
```

Download the data set:

```{python, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
url = "https://scholar.colorado.edu/downloads/zw12z650d"
fired = requests.get(url)
data_file = "fired.zip"
with open(data_file, 'wb') as f:
    f.write(fired.content)

# Unzip the file
with zipfile.ZipFile(data_file, 'r') as zip_ref:
    zip_ref.extractall()
```

Read it:

```{python, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
fired = gpd.read_file("fired_conus_ak_to_January_2022_gpkg_shp/conus_ak_to2022001_events.shp")
```

Plot fire duration as a function of ignition day:

```{python, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
plt.figure()
sns.scatterplot(data=fired, x='ig_day', y='event_dur')
sns.set_style('whitegrid')
plt.xlabel('Day')
plt.ylabel('Event duration (days)')
plt.show()
```
